Influence of Connectivity on Dung Beetle Communities

### 2025-06-06 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
added citations and refreshed bib

As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented(here, fragmentation). Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003] [@fletcher_is_2018]. However, as habitats are broken down community structures are significantly altered [@WOS:000419965700012]. This alteration of structure typically lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].

Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation[@haddad_corridor_2003]. By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy [@forman_general_1995]. Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity(here, trophic cascades).

By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes(experimental designs, maybe that one haddad). Past studies have measured changes in biodiversity for many different taxa [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected. Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services.

@@

Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al. 2024). The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013]
Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes. We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge. To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors. Additionally we asked are corridors benefiting any one functional trait over another?

@@

All dung beetles were counted and identified to species as described in @nemes_illustrated_2015-1 and @edmonds_taxonomic_2023. Fifteen individuals of each species with adequate captures were dried to equalibrium and weighed for biomass measurements. Voucher specimcens for each species will be deposited at the Florida State Collection of Arthropods.

@@

Biodiversity between patch types was compared using Hill numbers (Jost 2006). Hill numbers are metrics developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity. [delete: In our work, we use Hill numbers to explore the impacts of abundance in our diversity measurements]. They are an alternative to more specialized metrics such as _________, which are less intuitive for interpretation. [maybe a sentence on how / why hill numbers are easier to interpret?]  We looked at [compared] community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity). _[can you add a sentence about how these are calculated? by which I mean - are the calculations different than for richness, shannon, etc.?]_ Diversity numbers and species richness were calculated in R studio using package `hill`.Diversity numbers were calculated in R studio using package iNEXT (Chao et al. 2016). Bray-Curtis dissimiolarity values were calculated using package Vegan in R studio. Dung beetles were assigned traits by waste removal guild and habitat preference.

>>>>

As human disturbances continue to expand into natural landscapes, intact habitats are becoming increasingly fragmented(Diaz 2019). Like many ecological processes, fragmentation is a complex and multifaceted phenomenon bringing about many consequences which can be both positive and negative for ecosystems [@fahrig_effects_2003] [@fletcher_is_2018]. However, as habitats are broken down community structures are significantly altered [@WOS:000419965700012]. This alteration of structure typically lends to loss in biodiversity on a global scale and interruptions in ecosystem processes and functions [@haddad_habitat_2015].

Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation[@haddad_corridor_2003]. By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy [@forman_general_1995]. Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity(debinski and holt).

By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes[@fletcher_jr_landscape_2023; @resasco_contribution_2017]. Past studies have measured changes in biodiversity for many different taxa [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected. Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services(hevia 2017).

@@

Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al. 2024). The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013]
Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes. We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge[@tewksbury_corridors_2002]. To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors. Additionally we asked are corridors benefiting any one functional trait over another?

@@

All dung beetles were counted and identified to species as described in @nemes_illustrated_2015-1 and @edmonds_taxonomic_2023. Fifteen individuals of each species with adequate captures were dried to equilibrium and weighed for biomass measurements. Voucher specimens for each species will be deposited at the Florida State Collection of Arthropods.

@@

Biodiversity between patch types was compared using Hill numbers, a set of indeces developed with the goal of providing a unifying context for the quantification of the many ways we measure biodiversity [@jost_entropy_2006]. They are an alternative to more specialized metrics such as _________, which are less intuitive for interpretation. [maybe a sentence on how / why hill numbers are easier to interpret?]  We looked at [compared] community composition by increasing magnitudes of diversity components (qD) of 0D (species richness), 1D (Shannon entropy), and 2D (Simpson Diversity). _[can you add a sentence about how these are calculated? by which I mean - are the calculations different than for richness, shannon, etc.?]_ Diversity numbers and species richness were calculated in R studio using package `hill`.Diversity numbers were calculated in R studio using package iNEXT (Chao et al. 2016). Bray-Curtis dissimiolarity values were calculated using package Vegan in R studio. Dung beetles were assigned traits by waste removal guild and habitat preference.
